---
title: "The Effect of VAR on Football Penalties Received"
author: "Zak Vinter & Nick Power"
date: "12-07-2022"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'css/my-css.css']
    # self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---
exclude: true

```{r, setup, include = F}
library(knitr)
# Define colors
red_pink   = "#e64173"
turquoise  = "#20B2AA"
orange     = "#FFA500"
red        = "#fb6107"
blue       = "#3b3b9a"
green      = "#8bb174"
grey_light = "grey70"
grey_mid   = "grey50"
grey_dark  = "grey20"
purple     = "#6A5ACD"
slate      = "#314f4f"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  cache = F,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(knitr.table.format = "html")
```

---

# Background

**What is a Penalty?**

* Penalties are awarded to a team if a foul was committed against them within the opponents 18 yard box.

* Examples: handball, missed timed tackle, and clear obstruction of movement.


**What is VAR?**

* VAR stands for Video Assistant Referee.

* Facilitates pitch side review for on-field referees during the game.

* Designed to improve the accuracy of refereeing decisions throughout the sport.

---

# Natural Experiment


**Experiment Design**

* English Premier League, first country to implement policy (2019).

* Difference-in-difference style experiment.

* Scottish Premier League used as a comparison group.




---

# Data

* All penalties received in both leagues between 2013-2021, per club.

* 20 clubs per year in the English league and 12 for Scotland.

* Specific clubs are different year-on-year, with relegation and promotion.





---

# Exploratory Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(estimatr, modelsummary, tidyverse, ggplot2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

load("~/GitHub/Term_Project/R_Data/Joined/tidy_joined.Rdata")
load("~/GitHub/Term_Project/R_Data/England/tidy_eng.Rdata")
load("~/GitHub/Term_Project/R_Data/Scotland/tidy_scot.Rdata")
load("~/GitHub/Term_Project/R_Data/England/eng_pre.Rdata")
load("~/GitHub/Term_Project/R_Data/England/eng_post.Rdata")
load("~/GitHub/Term_Project/R_Data/Joined/Appended_Joined.Rdata")
load("~/GitHub/Term_Project/R_Data/England/Appended_Top_6.Rdata")
load("~/GitHub/Term_Project/R_Data/Econometrics/Reg_Data.Rdata")
load("~/GitHub/Term_Project/R_Data/Econometrics/Reg_Data_2.Rdata")
load("~/GitHub/Term_Project/R_Data/Econometrics/Reg_Data_3.Rdata")

```



```{r echo=FALSE, message=FALSE,warning=FALSE, fig.align = 'center'}
#line plot with points for average penalties per team in England
ggplot(data=tidy_eng,aes(x=years, y=pen_per_team)) + geom_point(color = "black", size = 2)+
  geom_line(color = "red", size = 0.8) +
  #set x axis limits and breaks for the year values in the data set
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  #set appropriate range for y values
  ylim(3,7) +
  #add title
  ggtitle("Average Penalties per Team in England") +
  #label the axis'
  xlab("Years") +
  ylab("Average Penalties (per team)") +
  #change the theme and the position of the title
  theme(plot.title = element_text(hjust = 0.5, color="black", size=14, face="bold"),
        #change theme of axis', tick markers
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(color="black", face="bold"), 
        axis.text.y = element_text(color="black", face="bold"),
  ) +
  #add line for the policies implementation date
  geom_vline(xintercept=2018.5, 
             color = "black", size=1) +
  #adding lines of best fit for both the pre and post policy time period
  geom_smooth(data = eng_pre, method = "lm", se = FALSE, col = "black") +
  geom_smooth(data = eng_post, method = "lm", se = FALSE, col = "black")




```

---
# Exporatory Analysis
```{r echo=FALSE, message=FALSE,warning=FALSE, fig.align = 'center'}

#line plot with points for average penalties per team in Scotland
ggplot(data=tidy_scot,aes(x=years, y=pen_per_team)) + geom_point(color = "black", size = 2)+
  geom_line(color = "blue", size = 0.8)+
  #set x axis limits and breaks for the year values in the data set
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  #set appropriate range for y values
  ylim(3,7) +
  #add title
  ggtitle("Average Penalties per Team in Scotland") +
  #label the axis'
  xlab("Years") +
  ylab("Average Penalties (per team)") +
  #change the theme and the position of the title
  theme(plot.title = element_text(hjust = 0.5, color="black", size=14, face="bold"),
        #change theme of axis', tick markers
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(color="black", face="bold"), 
        axis.text.y = element_text(color="black", face="bold"),
  ) +
  #line of best fit for full time period
  geom_smooth( method = "lm", se = FALSE, col = "black")



```


---

# Exploratory Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = 'center'}

#plot of two local regressions. Penalties per team for both Scotland and England
tidy_joined %>%
  ggplot(aes(years, pen_per_team, color = country)) +
  geom_smooth(size = 1.25) +
  #set x axis limits and breaks for the year values in the data set
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  #set appropriate range for y values
  ylim(3.5,6.5) +
  #add title
  ggtitle("Average Penalties per Team across Country") +
  #label the axis'
  xlab("Years") +
  ylab("Average Penalties (per team)") +
  #change the theme of the legend text
  guides(color = guide_legend(label.theme = element_text(color = c("black"), face ="bold"))) +
  #change the theme for the position of the title
  theme(plot.title = element_text(hjust = 0.5, color="black", size=14, face="bold"),
        #change theme of axis', tick markers
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(color="black", face="bold"), 
        axis.text.y = element_text(color="black", face="bold"),
        #remove legend title
        legend.title=element_blank()
        ) +
  #add line for the policies implementation date
  geom_vline(xintercept=2018.5, 
               color = "black", size=1)

```


---

# Exploratory Analysis
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = 'center'}



ggplot(data=Appended_Top_6, aes(x=Years, y=average_pens, color = Type)) + geom_line(size = 1) +
  geom_point(color = "black", size = 1) +
  #set x axis limits and breaks for the year values in the data set
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  #set appropriate range for y values
  ylim(2,12) +
  #add title
  ggtitle("Average Penalties per Rank in England") +
  #label the axis'
  xlab("Years") +
  ylab("Average Penalties (per Team)") +
  #change the theme of the legend text
  guides(color = guide_legend(label.theme = element_text(color = c("black"), face ="bold"))) +
  #change the theme and the position of the title
  theme(plot.title = element_text(hjust = 0.5, color="black", size=14, face="bold"),
        #change theme of axis', tick markers
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(color="black", face="bold"), 
        axis.text.y = element_text(color="black", face="bold"),
        #remove the legend title
        legend.title=element_blank()
  ) +
  #create line to represent the policy implementation 
  geom_vline(xintercept=2018.5, 
             color = "black", size=1)





```

---

# Regressions

**Main Model**

$$PenaltiesReceived_{ict}= β_0  + β_1 (Post_t * Treat_{ic}) + β_2  Post_t  + β_3  Treat_{ic} + ∈_{ict}$$

* $PenaltiesReceived_{ct}$ represents penalties received by individual club i in Country c in year t. 


* $β_1$ is the coefficient of interest, this measures the with-in country difference in penalties received before and after VAR is implemented, conditional on year fixed effects.

* $β_2$ represents country fixed effects, controlling for factors that are constant within country over-time. 

* $β_3$ represents time fixed effects, controlling for patterns over time that exist within the United Kingdom. 




---
# Addtional Regressions





$$Penalties_{ict}= β_0  + β_1 (Post_t * Top6_{ic}) + β_2  Post_t  + β_3 Top6_{ic} + ∈_{ict}$$


* Top six English Teams vs non-top six teams in England









$$Penalties_{ict}= β_0  + β_1 (Post_t * TreatTop6_{ic}) + β_2  Post_t  + β_3  TreatTop6_{ic} + ∈_{ict}$$

* Top teams in England vs top teams in Scotland



---

# Econometric Analysis


```{r echo=FALSE, message=FALSE, warning=FALSE}
reg_1_robust = lm_robust(PENALTIES_RECEIVED ~ Time + Area + Interaction, data = Reg_Data)

reg_2_robust = lm_robust(PENALTIES_RECEIVED ~ Time + Area + Interaction, data = Reg_Data_2)

reg_3_robust = lm_robust(PENALTIES_RECEIVED ~ Time + Area + Interaction, data = Reg_Data_3)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = 'center'}
modelsummary(list(
  "Reg_1" = reg_1_robust, 
  "Reg_2" = reg_2_robust, 
  "Reg_3" = reg_3_robust
), 
fmt = 2
)
```

---

# Problems for Causality


**Randomness of the policies implementation**


**Generalization of results**

* Style of play

* Financial Disparity


**Omitted Variables**

* Weather

* Pitch Quality



---

# Conclusions

**Findings**

* VAR has increased the number penalties received by clubs, however not statistically significantly.


* Policy has differential impacts across club size and ability.

**Further Research**

* Distinguish if is the video replay picking up real fouls that would have otherwise been missed without the technology, or is the technology causing excess penalties.











