---
title: "The Effect of VAR on Football Penalties Received"
author: "Zak Vinter & Nick Power"
date: "12-07-2022"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'css/my-css.css']
    # self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---
exclude: true

```{r, setup, include = F}
library(knitr)
# Define colors
red_pink   = "#e64173"
turquoise  = "#20B2AA"
orange     = "#FFA500"
red        = "#fb6107"
blue       = "#3b3b9a"
green      = "#8bb174"
grey_light = "grey70"
grey_mid   = "grey50"
grey_dark  = "grey20"
purple     = "#6A5ACD"
slate      = "#314f4f"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  cache = F,
  warning = F,
  message = F
)
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(knitr.table.format = "html")
```

---

# Background

**What is VAR?**

* VAR stands for Video Assistant Referee.

* Facilitates pitch side review for on-field referees during the game.

* Designed to improve the accuracy of refereeing decisions throughout the sport.







---

# Exploratory Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(estimatr, modelsummary, tidyverse, ggplot2)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

load("~/documents/GitHub/Term_Project/R_Data/Joined/tidy_joined.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/England/tidy_eng.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/Scotland/tidy_scot.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/England/eng_pre.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/England/eng_post.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/Joined/Appended_Joined.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/England/Appended_Top_6.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/Econometrics/Reg_Data.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/Econometrics/Reg_Data_2.Rdata")
load("~/documents/GitHub/Term_Project/R_Data/Econometrics/Reg_Data_3.Rdata")

```



```{r echo=FALSE, message=FALSE,warning=FALSE, fig.align = 'center'}
#line plot with points for average penalties per team in England
ggplot(data=tidy_eng,aes(x=years, y=pen_per_team)) + geom_point(color = "black", size = 2)+
  geom_line(color = "red", size = 0.8) +
  #set x axis limits and breaks for the year values in the data set
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  #set appropriate range for y values
  ylim(3,7) +
  #add title
  ggtitle("Average Penalties per Team in England") +
  #label the axis'
  xlab("Years") +
  ylab("Average Penalties (per team)") +
  #change the theme and the position of the title
  theme(plot.title = element_text(hjust = 0.5, color="black", size=14, face="bold"),
        #change theme of axis', tick markers
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        axis.text.x = element_text(color="black", face="bold"), 
        axis.text.y = element_text(color="black", face="bold"),
  ) +
  #add line for the policies implementation date
  geom_vline(xintercept=2018.5, 
             color = "black", size=1) +
  #adding lines of best fit for both the pre and post policy time period
  geom_smooth(data = eng_pre, method = "lm", se = FALSE, col = "black") +
  geom_smooth(data = eng_post, method = "lm", se = FALSE, col = "black")




```



---
class: inverse, middle
name: bigdata

# Tools for big data


---

# Cloud computing

**Set up a virtual machine (VM) on a cloud computing platform.**
* Easy and cheap. (Much more so than setting up your own server!)
* Get billed for exactly as much power, memory, and time as you need.
* For most purposes, you won't use more than a few dollars per month.

**Platforms:**
* [Google Cloud Compute Engine](https://cloud.google.com/compute/)
* [Amazon Web Services](https://aws.amazon.com/)
* [Posit Cloud](https://posit.cloud/) (affiliated with RStudio).

**Tutorials** by Grant McDermott: [Part 1](https://raw.githack.com/uo-ec510-2020-spring/lectures/master/13-gce-i/13-gce-i.html), [Part 2](https://raw.githack.com/uo-ec510-2020-spring/lectures/master/14-gce-ii/14-gce-ii.html).


