---
title: "Untitled"
author: "Zak Vinter"
date: "2022-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = 'center'}
# vertical graphs stacked on top of each other

# create datasets with just england and scotland totals.
england_grouped<- Appended_Joined|> 
  filter(COUNTRY=="England")


scot_grouped<- Appended_Joined|> 
  filter(COUNTRY=="Scotland") 


#create England basic plot
hat<-ggplot(data = england_grouped,aes(x=YEAR,y=TOTAL))+ scale_x_continuous(limits= c(2013,2021),
                                                                        
                                  breaks=c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
                      scale_y_continuous(limits=c(0,150), 
                                         minor_breaks= NULL, 
                                         expand=c(0,1))+geom_smooth(span=0.65, se=FALSE) +ggtitle("England")+
                                          theme(plot.title = element_text(hjust=.5))




#create basic scotland plot

bat<-ggplot(data=scot_grouped,aes(x=YEAR, y=TOTAL)) + 
  scale_x_continuous(limits = c(2013,2021), 
                     breaks = c(2013,2014,2015,2016,2017,2018,2019,2020,2021))+
  scale_y_continuous( limits=c(0,150),
                      minor_breaks = NULL, 
                      expand = c(0,1)) +geom_smooth(span=0.65,se=FALSE) +ggtitle("Scotland")+
  theme(plot.title=element_text(hjust = .5))




# load the cowplot package to allow combination
library(tidyverse)
if (!require("dslabs")) install.packages("dslabs")
library(dslabs)
  

library(cowplot)



# the title of the graph
title<- ggdraw()+ draw_label("Penalties by Year for England & Scotland",
              fontface = 'bold', x=0, hjust=-.11) + 
  theme(plot.margin = margin(0,0,0,7))




# the plots combined together.
combined_plots<-cowplot::plot_grid(hat,bat, labels="",label_size = 6, ncol = 1)
         

# the plots combined with the main title.         
plot_grid(title,combined_plots, ncol=1, rel_heights = c(0.1,1))






```

The final graph is just an overall illustration of the Total number of penalties per country by year, not by team. This was chosen as the y-axis instead of penalties per team because it shows a relationship that may be a bit easier to interpret. Since it is Just the total number of penalties throughout both leagues per year it can have a little less variation as the averages per team per year may show some more potential outlying behavior. This graph nicely illustrates the trends on top of each other. It shows how the overall penalties by year is on a larger scale than for the average per team per year, which can help get a birds-eye view of the overall situation. It may be slight in appearance, but the slight trend of total number of penalties in England seems to make a marked increase compared to Scotland in 2018, right when the policy takes place.




